\documentclass[../main.tex]{subfiles}
\graphicspath{{\subfix{./figures/}}}
\begin{document}
The purpose of this chapter is to give the necessary background information
in rational homotopy theory for our proof of the rational homotopy groups
of spheres. We shall give the definition of a rational homotopy
equivalence and state some elementary results concerning these
equivalences. Much of what we do here is taken from Chapter 4 of \cite[]{Ber12}.

\section{Rational homotopy equivalence}
\begin{definition}[Rational homotopy equivalence]
    Let \( f:X\rightarrow Y \) be a map. We say that \( f \) is a
    \emph{rational homotopy equivalence} if the induced map on
    rationalized homotopy groups, \( \pi_*(f)\otimes\Q:\pi_*(X)\otimes
    \Q\rightarrow\pi_*(Y)\otimes\Q\), is an isomorphism.
\end{definition}
This weakens the criterion of weak homotopy equivalence which requires
that \( \pi_*(f):\pi_*(X)\rightarrow\pi_*(Y) \) be an isomorphism. Thus,
both every homotopic and weakly homotopic spaces are rationally homotopic.
It turns out, due to a result of Serre, that the above definition is
equivalent to requiring that the induced map in homology with rational
coefficients is an isomorphism.
\begin{proposition}
    Let \( f:X \rightarrow Y \) be a map between spaces. Then, \( f \)
    is a rational homotopy equivalence if and only if the induced
    map \( f_*:H_*(X;\Q)\rightarrow H_*(Y;\Q) \) is an isomorphism.
\end{proposition}
\begin{proof}
    See \cite[]{Ser53}.
\end{proof}
We also have an analogous result for cohomology which requires the notion
of a \emph{uniquely divisible group}.
\begin{definition}[Uniquely divisible group]
    An abelian group \( G \) is said to be \emph{divisble} if for all
    \( g \in G \) and all positive integers \( n \) there exists \( y
    \in G\) such that
    \( ny=g \). The group is uniquely divisible if \( y \) is unique.
\end{definition}
\begin{proposition}
   \label{prop:rat-cohomology}
   For any map \( f:X \rightarrow Y \) we have that it is a rational
   homotopy equivalence if and only if for any uniquely divisible
   abelian group \( G \) the induced map \( f^*:H^*(Y;G)\rightarrow 
   H^*(X;G)\) is an isomorphism.
\end{proposition}
\begin{proof}
    See \cite[]{Ber12}.
\end{proof}

\end{document}
